### 2024-02-07 test.qmd
Update test.qmd

---
title: "test"
format: html
---

#data cleaning -> a column called year will be produced in the data 
#(you can change the names of the data set)
data_strikes_year <- dataset_strikes
data_strikes_year$Year <- substr(data_strikes_year$date, 1, 4)