### 2024-02-07 test.qmd
Update test.qmd

---
title: "test"
format: html
---
#install.packages(haven)
library(haven)
library(ggplot2)
library(tidyverse)
library(dplyr)

```{r}
dataset_events <- read_dta("dataset_events.dta")

dataset_holiday <- read_dta("dataset_holiday.dta")

dataset_strikes <- read_dta("dataset_strikes.dta")

dataset_weather <- read_dta("dataset_weather.dta")

distancematrix <- read_dta("distancematrix.dta")

pollstations_id <- read_dta("pollstations_id.dta")

stadtkat_id <- read_dta("stadtkat_id.dta")

```

#data cleaning -> a column called year will be produced in the data 
#(you can change the names of the data set)
data_strikes_year <- dataset_strikes
data_strikes_year$Year <- substr(data_strikes_year$date, 1, 4)